\section{Conclusion:}
As a mostly script-focused article, there is really no need for a conclusion, but I like writing them just the same. This article has two purposes to it, and I believe it has satisfied both. The first was to document the CPU Thermal scripts I have developed, so those interested can learn more about my reasoning behind them and how they function. The second was to give me an excuse to review and overhaul the scripts, a process that led to improvements to design and the addition of features. The design improvements should make it easier to add tests in the future while the new pulse system is reliable, unlike the original. It also gave me an idea for how to control the length of the pulse, not just the length of the pause, but I do not feel this is necessary except for running very controlled tests.

The examination of Prime95's behavior was interesting, as I was not expecting it to be as variable as it was. While that does weaken its usefulness for tests like these scripts were designed for, it could still be of interest to some and provides another CPU load option for everyone. The consistency of the Cinebench benchmarks makes them a better load for these scripts.

The new pulsing feature made it possible to examine how AMD's boosting methods work, and I feel confirmed they are opportunistic, rather than deterministic, an important distinction. A deterministic boost will simply increase the frequency by some amount based on certain conditions, such as the temperature and power limits. An opportunistic boost will still be controlled by conditions, but will push the frequency as far as it can when loaded, rather than a fixed amount. This can lead to some interesting results, such as the spikes we saw at the end of pauses or when the load changes. It is also opportunistic boosting that has led some to claim overclocking is a dying art, but that is a separate topic with a reasonable amount of depth to it.

I am not sure what may be the future of these scripts, but I am glad I originally developed them, and quite happy I have taken the time to overhaul them now.
